Motional Quantum Ground-State Cooing of a Single Sodium Atom

### 2017-06-01 paper.tex

Single neutral atoms trapped in optical tweezers provide a promising system
for various applications including quantum information, quantum chemistry
and quantum simulation of many-body systems.
The interaction between different sites can be controlled by creating Rydberg excitation
or by creating dipolar molecules from two trapped atoms.
Combined with the ability to move, detect and manipulate individual sites,
arrays of optical tweezers can be used to prepare complex interacting system
with high fidelity.
In order to achieve long coherence time and full quantum control of the system needed for
these applications,
it is usually necessary to control the thermal motion of the atoms by cooling to the
three dimensional motional ground state in the optical tweezer,
which has already been demonstrated with heavy alkali atoms including rubidium and cesium.
However, some of the important properties of previous experiments,
for example low initial temperature and small Lamb-Dicke parameter,
may not be easily achievable for other systems and
the versatility of this approach requires being able to perform ground motional state cooling
in the more general case. In this letter, we present our work on cooling single sodium atoms
trapped in optical tweezers to the motional ground state with Raman sideband cooling.
Despite having a large Lamb-Dicke parameter and high initial temperature,
by utilizing several new cooling techniques and a well optimized cooling sequence,
we are able to achieve a ground state probability of $80\%$.

>>>>

Single neutral atoms trapped in optical tweezers provide a promising system
for various applications including quantum information, quantum chemistry
and quantum simulation of many-body systems.
Long-range interactions between the sites can be achieved by using Rydberg atoms, polar molecules, or dipolar atoms.
Polar molecules offer the additional advantage of a high degree of tunability, including the shape, range, and strength of the interaction.
Arrays of optical tweezers offer the ability to move, detect, and manipulate individual sites, andcan be used to prepare complex interacting systems
with high fidelity.
In order to achieve long coherence time and full quantum control of the system needed for
these applications,
it is usually necessary to control the thermal motion of the atoms by cooling to the
three dimensional motional ground state in the optical tweezer,
which has already been demonstrated with heavy alkali atoms including rubidium and cesium.
However, some of the important properties of previous experiments,
for example low initial temperature and small Lamb-Dicke parameter,
may not be easily achievable for other systems, such as directly laser-cooled polar molecules, or atoms with challenging sub-Doppler cooling such as Na, K, and Li.
To extend the versatility of this approach, we must be able to apply it in a more general case. 
In this letter, we cool single sodium atoms
trapped in optical tweezers to the motional ground state with Raman sideband cooling.
Despite having a large Lamb-Dicke parameter and high initial temperature, we are able to achieve a ground state probability of $XX\%$
by utilizing several new cooling techniques and a carefully optimized cooling sequence.
